### 2024-05-26 thesis/input_files/data_exploratory.tex
corrected all comments

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_total_load.png}
    \caption{Zonal price versus total load GEFCom2014}
    \label{fig:gefcom_zonal_price_vs_total_load}
\end{figure}

\subsection{GEFCom2014 load track}
In the load track, contestants were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Any data from the various task folders are shifted by one month between each other.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_\-V2.csv file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
The pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the historical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally, figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
These plots motivate the selection of weather temperature as one of the predictors for the load.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_load_historical.png}
    \caption{Historical load GEFCom2014}
    \label{fig:gefcom_load_historical}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_w_avg_historical.png}
    \caption{Historical weather GEFCom2014}
    \label{fig:gefcom_w_avg_historical}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_load_vs_w_avg.png}
    \caption{Load versus weather GEFCom2014}
    \label{fig:gefcom_load_vs_w_avg}
\end{figure}

% \subsection{GEFCom ETL}
\subsubsection{GEFCom ETL}
The provided data is raw, hence the need for a standardized extract transform and load pipeline. 
Firstly, we drop all Nans and zeros. Notice, the provided data for the challenge has a timestamp column not coherent to any time format. Thus, we first pass the data through a method extracting time information: hour, day, month and year. This info is then combined into a datetime object, we selected the ISO format.
Next, since the data does not provide the location for the target utility, the best we can do is selecting the average of weather temperatures as predictor variables. This independent variable is named w avg.
The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming convention and formatting used in the last track of the challenge. The Python ETL pipeline has been then automated to loop every task folder through Bash scripting.

\section{Germany Switzerland Austria data}
Next, we identified various data sources for Germany, Switzerland and Austria in order to further
investigate the techniques studied in this thesis work.
Taking inspiration from the GEFCom2014 format, we firstly created a load track for Germany and Switzerland. 
The source of data has been the \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} website. It provides data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations. Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{(ENTSOE)} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
\\
In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see section \ref{ch:prob}.
The target dependent variable is the national load while as predictors we took
the weather temperature, the wind speed, the date expressed in terms of the three attributes day, month and hour, a categorical variable for the day of the week and a binary categorical variable accounting for the holiday effect.

In the second experiment we compared various kernels in the context of load forecasting.
For this study we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}. This data consists of various metereological attributes relevant for modelling electricity demand \cite{Formayer2023}. The dataset is made up of historical data until end of 2020, while from 2021 onwards the data consists of forecasts made by a metereological model developed by the European Centre for Medium-Range Weather Forecasts (ECMWF). Notice, the forecasted data is splitted in two files according to the scenarios RCP4.5 and RCP8.5. In our experiments we considered the former. 
In this setting the independent variables available are 
direct irradiation, global radiation, power from reservoir plants, power from run of the river plants (ROR), weather temperature and wind potential.

\subsubsection{ETL}
From energy charts we downloaded the historical load along with the weather temperatures and wind speeds forecasts. Notice, the data request to the environmental variables at hourly frequency for a whole year is broken. Therefore, we downloaded each month separately and then transformed them into a single csv for the whole year.

For what concerns the data from SECURES-Met we first identified the data interesting to us. Next, we formatted all the csv to a coherent format, gave columns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv\textsubscript{s} together.

@@

\subsubsection{EDA}
Following are reported the EDA plots alongside with insights that can be extracted from them. We choose to focus the following analysis on the Swiss data only, given that the plots for the German data are pretty similar.
To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}. As expected, electricity load exhibits a seasonal pattern with highs in the winter and lows during the summer season.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
    \caption{Switzerland historical load (2021)}
    \label{fig:CH_historical_load_2021}
\end{figure}
Next, we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed. From these plots we can clearly see there exists a relationship that ties wind speeds and temperatures to load.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_load_vs_temperature_2021.png}
    \caption{Load vs temperature Switzerland (2021)}
    \label{fig:CH_load_vs_temperature_2021}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_load_vs_wind_speed_2021.png}
    \caption{Load vs wind speed Switzerland (2021)}
    \label{fig:CH_load_vs_wind_speed_2021}
\end{figure}

Following, we used box plots in order to visualize the relations between categorical independent variables and the dependent one. Such visualisation are reported in figure \ref{fig:CH_day_of_week_boxplot_2021} and figure \ref{fig:CH_is_holiday_boxplot_2021}. We can see that weekdays and workdays involve a higher load and possess a wider range.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_day_of_week_boxplot_2021.png}
    \caption{Load versus day of week boxplot Switzerland (2021)}
    \label{fig:CH_day_of_week_boxplot_2021}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_is_holiday_boxplot_2021.png}
    \caption{Load versus day type boxplot Switzerland (2021)}
    \label{fig:CH_is_holiday_boxplot_2021}
\end{figure}

We conclude this EDA by reporting the heatmap of independent variables alongside with the load in figure \ref{fig:CH_heatmap_2021}. Temperature, wind speed and day of week show a meaningful correlation. This is a clear indication of their suitability as predictors for forecasting the load.

>>>>

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_total_load.png}
    \caption{Zonal price versus total load of GEFCom2014}
    \label{fig:gefcom_zonal_price_vs_total_load}
\end{figure}

\subsection{GEFCom2014 load track}
In the load track, contestants were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Any data from the various task folders are shifted by one month between each other.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_\-V2.csv file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
The pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the historical load while Figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally, Figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
These plots motivate the selection of weather temperature as one of the predictors for the load.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_load_historical.png}
    \caption{Historical load of GEFCom2014}
    \label{fig:gefcom_load_historical}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_w_avg_historical.png}
    \caption{Historical weather of GEFCom2014}
    \label{fig:gefcom_w_avg_historical}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_load_vs_w_avg.png}
    \caption{Load versus weather of GEFCom2014}
    \label{fig:gefcom_load_vs_w_avg}
\end{figure}

% \subsection{GEFCom ETL}
\subsubsection{GEFCom ETL}
The provided data is raw, hence the need for a standardized extract transform and load pipeline. 
Firstly, we drop all Nans and zeros. The provided data for the challenge has a timestamp column not coherent to any time format. Thus, we first pass the data through a method extracting time information: hour, day, month and year. This info is then combined into a datetime object, we selected the ISO format.
Next, since the data does not provide the location for the target utility, the best we can do is selecting the average of weather temperatures as predictor variables. This independent variable is named w avg.
The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming conventions and formatting used in the last track of the challenge. The Python ETL pipeline has since been automated to loop every task folder through Bash scripting.

\section{Germany Switzerland Austria data}
Next, we identified various data sources for Germany, Switzerland and Austria in order to further
investigate the techniques studied in this thesis.
Taking inspiration from the GEFCom2014 format, we firstly created a load track for Germany and Switzerland. 
The source of data has been the \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} website. It provides data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations. This data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{(ENTSOE)} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
\\
In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see Section \ref{ch:prob}.
The target dependent variable is the national load while as predictors we took
the weather temperature, the wind speed, the date expressed in terms of the three attributes day, month and hour, a categorical variable for the day of the week and a binary categorical variable accounting for the holiday effect.

In the second experiment we compared various kernels in the context of load forecasting.
For this study, we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}. This data consists of various metereological attributes relevant for modelling electricity demand \cite{Formayer2023}. The dataset is made up of historical data until end of 2020, while from 2021 onwards the data consists of forecasts made by a metereological model developed by the European Centre for Medium-Range Weather Forecasts (ECMWF). The forecasted data is split in two files according to the scenarios RCP4.5 and RCP8.5. In our experiments, we considered the former. 
In this setting the independent variables available are 
direct irradiation, global radiation, power from reservoir plants, power from run of the river plants (ROR), weather temperature and wind potential.

\subsubsection{ETL}
From energy charts we downloaded the historical load along with the weather temperatures and wind speeds forecasts. The data request to the environmental variables at hourly frequency for a whole year is broken. Therefore, we downloaded each month separately and then transformed them into a single csv for the whole year.

For what concerns the data from SECURES-Met we first identified the data interesting to us. Next, we formatted all the csv to a coherent format, gave columns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally merged all single csv\textsubscript{s} together.

@@

\subsubsection{EDA}
Following we report the EDA plots alongside with insights that can be extracted from them. We choose to focus the following analysis on the Swiss data only, given that the plots for the German data are pretty similar.
To get started we visualise the historical load data in Figure \ref{fig:CH_historical_load_2021}. As expected, electricity load exhibits a seasonal pattern with highs in the winter and lows during the summer season.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
    \caption{Switzerland historical load (2021)}
    \label{fig:CH_historical_load_2021}
\end{figure}
Next, we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while Figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed. From these plots, we can clearly see there exists a relationship that ties wind speeds and temperatures to load.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_load_vs_temperature_2021.png}
    \caption{Load versus temperature for Switzerland (2021)}
    \label{fig:CH_load_vs_temperature_2021}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_load_vs_wind_speed_2021.png}
    \caption{Load versus wind speed for Switzerland (2021)}
    \label{fig:CH_load_vs_wind_speed_2021}
\end{figure}

Following, we use box plots in order to visualize the relations between categorical independent variables and the dependent one. Such visualisation are reported in Figure \ref{fig:CH_day_of_week_boxplot_2021} and Figure \ref{fig:CH_is_holiday_boxplot_2021}. We can see that weekdays and workdays involve a higher load and possess a wider range.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_day_of_week_boxplot_2021.png}
    \caption{Load versus day of week boxplot for Switzerland (2021)}
    \label{fig:CH_day_of_week_boxplot_2021}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_is_holiday_boxplot_2021.png}
    \caption{Load versus day type boxplot for Switzerland (2021)}
    \label{fig:CH_is_holiday_boxplot_2021}
\end{figure}

We conclude this EDA by reporting the heatmap of independent variables alongside with the load in Figure \ref{fig:CH_heatmap_2021}. Temperature, wind speed and day of week show a meaningful correlation. This is a clear indication of their suitability as predictors for forecasting the load.

### 2024-05-26 thesis/input_files/implementation.tex
corrected all comments

This section is intended to explain and aid for reproducibility studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
\\
For the list of Python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
% - indicate computer specifics
All experiments have been carried out on a 3.2 GHz 16GB Apple M1 Pro.

@@

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
In our application we specified as hyperparameters the length of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.

@@

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented Python open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point method. This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the $Q$ matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator. To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.

>>>>

This section is intended to explain and aid for reproducibility studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
\\
For the list of Python packages needed, see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
% - indicate computer specifics
All experiments have been carried out on a 3.2 GHz 16GB Apple M1 Pro.

@@

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
The Tbats implementation is available at \href{https://github.com/intive-DataScience/tbats}{https://github.com/intive-Data-Science/tbatsx}.
In our application we specified as hyperparameters the length of seasons, that is, 24 for the daily seasonality and 168 for the weekly seasonality.

@@

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented Python open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularisation term $C$, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point method. This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the $Q$ matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a single matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with in built scikit-learn methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}. Functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator. To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.

### 2024-05-26 thesis/input_files/literaturereview.tex
corrected all comments

\section{Kernel methods literature review}
%INTRO KERNEL METHODS E ENERGY PRICE MODELLING
Kernel methods are a class of algorithms for pattern analysis.
With kernel methods we are able to apply linear methods with predictors in a high dimensional space, without having to explicitly evaluate the involved dot products of the features.
Throughout this thesis, we will address the performance of kernel methods in the context of EF.
\\
%- Kernel history evolution.
Their name comes from the German word kern, which translates to core in english. Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he introduced the term "definite kernels". Following, Hilbert's and Schmidt's work, \cite{schmidt} lead to the introduction of a new space, the Hilbert space.
In 1909, James Mercer improved Hilbert's work by proposing his theorem \cite{mercer}. This theorem underlies the power of kernel methods, that is the kernel trick.
In 1938, Schoenberg \cite{schoenberg} developed the mathematical results that allow us to find the kernel associated to a specific feature space metric.
In 1941, Kolmogorov \cite{kolmogorov} carried out stuides on representing kernels in linear spaces.
In 1950, Aronszajn \cite{aronszajn} published the first work on reproducing kernel hilbert spaces (RKHS); developing the general method for representing kernels in linear spaces.
In 1964, Aizerman \cite{aizerman} further improved the theory of RKHS.
It was in the 1990s that the theory of kernel methods got popular, particularly in the field of machine learning. Kernels have been used in various different tasks such as support vector machines (SVM) \cite{vapnik1} \cite{vapnik2}, Gaussian process classifiers \cite{williams}, spline methods \cite{wahba}, neural networks \cite{poggio} and principal component analysis \cite{pca_scholkopf}.
Nevertheless kernel methods received very little attention in the specific setting of EF literature.
\\
The kernel theory needed for this thesis work is covered in Section \ref{kernel_theory}. 
For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review} and \cite{shawe2004kernel}.
\\
Kanagawa et Fukumizu introduces to the concept of kernel mean embedding \cite{pmlr}. Muandet et al. \cite{Muandet_2017} surveys established results and new advances in the theory of Hilbert space distribution embeddings. 
It has to be said that, computing and storing such embeddings becomes prohibitive for large scale settings. Rudi et al. \cite{2022nystrom} proposes an efficient approximation procedure based on the Nyström method \cite{nystrom}, providing also an upper bound for the approximation error.
Another example of the applicability of kernel methods is \cite{supersamples}. In that paper, Smola et al.\ kernelize the herding algorithm introduced in \cite{welling2009herding,welling2009herding,Welling2010}. The original herding algorithm converts empirical moments into a sequence of pseudo-samples approximating the probability density function (PDF) of the observed random variable. Its weakness is that it can only take a finite number of features and this limits its expressive power.
Smola et al.\ overcome this by translating the algorithm into the framework of kernels.

>>>>

\section{Kernel methods literature review}
%INTRO KERNEL METHODS E ENERGY PRICE MODELLING
Kernel methods are a class of algorithms for pattern analysis.
With kernel methods we are able to apply linear methods with predictors in a high dimensional space, without having to explicitly evaluate the involved dot products of the features.
Throughout this thesis, we will address the performance of kernel methods in the context of EF.
\\
%- Kernel history evolution.
Their name comes from the German word kern, which translates to core in english. Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he introduced the term "definite kernels". Following, Hilbert's and Schmidt's work, \cite{schmidt} lead to the introduction of a new space, the Hilbert space.
In 1909, James Mercer improved Hilbert's work by proposing his theorem \cite{mercer}. This theorem underlies the power of kernel methods, that is the kernel trick.
In 1938, Schoenberg \cite{schoenberg} developed the mathematical results that allow us to find the kernel associated to a specific feature space metric.
In 1941, Kolmogorov \cite{kolmogorov} carried out stuides on representing kernels in linear spaces.
In 1950, Aronszajn \cite{aronszajn} published the first work on reproducing kernel Hilbert spaces (RKHS); developing the general method for representing kernels in linear spaces.
In 1964, Aizerman \cite{aizerman} further improved the theory of RKHS.
It was in the 1990s that the theory of kernel methods got popular, particularly in the field of machine learning. Kernels have been used in various different tasks such as support vector machines (SVM) \cite{vapnik1} \cite{vapnik2}, Gaussian process classifiers \cite{williams}, spline methods \cite{wahba}, neural networks \cite{poggio} and principal component analysis \cite{pca_scholkopf}.
Nevertheless kernel methods received very little attention in the specific setting of EF literature.
\\
The kernel theory needed for this thesis work is covered in Section \ref{kernel_theory}. 
For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review} and \cite{shawe2004kernel}.
\\
Kanagawa et Fukumizu introduces to the concept of kernel mean embedding \cite{pmlr}. Muandet et al. \cite{Muandet_2017} surveys established results and new advances in the theory of Hilbert space distribution embeddings. 
It has to be said that, computing and storing such embeddings becomes prohibitive for large scale settings. Rudi et al. \cite{2022nystrom} proposes an efficient approximation procedure based on the Nyström method \cite{nystrom}, providing also an upper bound for the approximation error.
Another example of the applicability of kernel methods is \cite{supersamples}. In that paper, Smola et al.\ kernelize the herding algorithm introduced in \cite{welling2009herding,welling2009herding,Welling2010}. The original herding algorithm converts empirical moments into a sequence of pseudo-samples approximating the probability density function (PDF) of the observed random variable. Its weakness is that it can only take a finite number of features and this limits its expressive power.
Smola et al.\ overcome this by translating the algorithm into the framework of kernels.
